---
title: "assignment_2"
author: "Hermes Martinez"
date: "10/5/2017"
output: html_document
---

## SETUP
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("functions.R")
```

## EXERCISE 1

Exercice 1a.
```{r}
print(sum_column(iris, "Sepal.Length"))
print(sum_column(iris, "Species"))
print(sum_column(warpbreaks, "breaks"))
```

Exercice  1b.
```{r}
print(my_sum(iris$Sepal.Length))
print(my_sum(iris$Species))
print(my_sum(warpbreaks$breaks))
```

Exercice 1c.
```{r}
print(sum_divided_by(iris$Sepal.Length, 12))
print(sum_divided_by(iris$Species, 22))
print(sum_divided_by(iris$Sepal.Length, "Not numeric"))
print(sum_divided_by(warpbreaks$breaks, -12))
```

Exercice 1d.
```{r}
print(my_mean(iris$Sepal.Length))
print(my_mean(iris$Species))
print(my_mean(warpbreaks$breaks))
```

## EXERCISE 2

Exercice 2a.
```{r}
print(grouped_violin_plot(iris, "Sepal.Length", "Species"))
```

Exercice 2b. 
```{r}
p <- grouped_violin_plot(iris, "Sepal.Length", "Species")
# YOUR CODE HERE: Change the colour scheme for the interior of the three violin plots
#                 to anything else at all.
p <- p + ggplot2::scale_fill_brewer(palette="black")
# YOUR CODE HERE: Add a main title that says "Iris data".
p <- p + ggplot2::ggtitle("Iris Data")
print(p)
```

## EXERCISE 3

Exercice 3a.
```{r}
difference_in_medians(iris, "Sepal.Width", "Species", "versicolor", "virginica")
difference_in_medians(iris, "Sepal.Width", "Species", "virginica", "virginica")
```

Exercice 3b.
```{r}
iris$Sepal.Width[1:10]
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
randomize(iris, "Sepal.Width")$Sepal.Width[1:10]
randomize(iris, "Species")$Species[1:10]
randomize(iris, "Species")$Sepal.Width[1:10]
set.seed(previous_seed)
```

Exercice 3c.
```{r}
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
ptest_1 <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor",
                                 "virginica", difference_in_medians, n_samples=10)
ptest_2 <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor",
                                 "virginica", difference_in_medians, n_samples=10)
ptest_3 <- permutation_twogroups(randomize(iris, "Sepal.Width"), "Sepal.Width",
                                 "Species", "versicolor", "virginica",
                                 difference_in_medians, n_samples=10)
set.seed(previous_seed)
print(ptest_1)
print(ptest_2)
print(ptest_3)
print(ptest_3$observed)
print(ptest_3[["observed"]])
```

Exercice d. Does it matter for your permutation test whether you permute var or grouping_var? Why or why not?
Yes, it does matter if you change var to grouping_var for instance, the outcome changes from "Sepal.Width" (expected) to an unexpected permutation with the Species group.

Exercice 3e.
```{r}
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
ptest <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor",
                              "virginica", difference_in_medians)
set.seed(previous_seed)


```


```{r}
ptest_d <- tibble::as_tibble(ptest["permuted"])
p <- ggplot2::ggplot(ptest_d, ggplot2::aes_string(x="permuted"))
p <- p + ggplot2::geom_histogram(fill="blue", colour="black")
p <- p + ggplot2::ggtitle("Exercice 3e")
p <- p + ggplot2::geom_vline(ggplot2::aes(xintercept=ptest[["observed"]]),lwd=0.7)
print(p)
  
```

Exercice 3f.
```{r, cache=T}
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
ptest_new <- permutation_twogroups(iris, "Sepal.Width", "Species", "versicolor",
                                   "virginica", new_test_statistic)
set.seed(previous_seed)
```
```{r}
ptest_n <- tibble::as_tibble(ptest_new["permuted"])
p2 <- ggplot2::ggplot(ptest_n, ggplot2::aes_string(x="permuted")) +
  ggplot2::geom_histogram(colour="black", fill="yellow") +
  ggplot2::geom_vline(ggplot2::aes(xintercept=ptest_new[["observed"]]), lwd=0.6, color="black") +
print(p2)
```

Exercice 3g.

```{r}
pvalue_r <- permutation_pvalue_right(ptest)
pvalue_l <- permutation_pvalue_left(ptest)
print(pvalue_r)
print(pvalue_l)
```
